---
title: "Credit Risk Analysis"
author: "Matilda Bosire"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Reading in the Data

```{r}

library(readr)
credit_data <- read_csv("F:/Strathmore Coursework/Applied Analytics/credit applied analytics.csv")
head(credit_data)

```

# Data Preprocessing

## Overview of Data

```{r}
class(credit_data)
```

```{r}
summary(credit_data)
```
## Sample Data Plots

```{r}

par(mfrow = c(2,2))

# frequency plots of numeric variables 
hist(credit_data$default, main = "Plot of Debtors Default in Credit History", xlab = "Default")
hist(credit_data$age, main = "Plot of Ages of Debtors in Credit History", xlab = "Age")

# frequency plots of character variables 
plot(as.factor(credit_data$job), main = "Plot of Job Groups of Debtors in Credit History", xlab = "Job Groups")
plot(as.factor(credit_data$housing), main = "Plot of Housing of Debtors in Credit History", xlab = "Housing")

```

## Character to Numeric Data

```{r}

# Factor out numeric columns
num_data <- credit_data[ , 1:8]
head(num_data )

```

```{r}

# function to change categorical character data to numeric data

checking_balance <- factor(credit_data$checking_balance, ordered = TRUE)
head(checking_balance)

head(as.numeric(checking_balance))

```

```{r}

# generalizing the function to apply to all columns

char_cols <- credit_data[ , 9:ncol(credit_data)]

fact_function <- function(x){
  return(as.numeric(factor(x, ordered = TRUE)))
}

char_data <- lapply(char_cols, fact_function)
char_data <- as.data.frame(char_data, col.names = colnames(char_cols))
head(char_data)

```

```{r}
# the final dataset

data <- cbind.data.frame(num_data, char_data)
head(data)
```

# Fitting a Linear Regression Model

```{r}

linear_model <- lm(data$default~., data = data)

summary(linear_model)
```

### Checking for Homoscedasticity

```{r}
par(mfrow = c(2,2))
plot(linear_model)
```

# Fitting a Logistic Regression Model 

This iks given that our response variable is binary where 1 represents default event, while 2 (now changed to 0) represents no default. 

## Train - Test Split

## Manually (80 - 20)

```{r}

# we have 1000 data points (rows) on 21 variables (columns)
# 80% of this is 800 rows

train_data <- data[1:800, ]
test_data <- data[801:1000, ]
```

## Fitting a linear model on the train data

```{r}
training_logistic_model <- glm(train_data$default~., data = train_data, family = "binomial")

summary(training_logistic_model)
```

## Automatic Split

Much better and functional

```{r}
# install.packages("caTools") # for logistic regression
# install.packages("ROCR") # for model evaluation

library(caTools)
library(ROCR)
```


```{r}
split_data <- sample.split(data, SplitRatio = 0.8)

# split ration is 80:20

head(split_data)
```
```{r}
traindata <- subset(data, split_data == "TRUE")  # picks datapoints that return TRUE
testdata <- subset(data, split_data == "FALSE")
```


### Fitting the Logistic Model

```{r}

# Logistic Model on the Training model
logistic_model <- glm(traindata$default~., data = traindata, family = "binomial")
summary(logistic_model)

```

```{r}

train_predict <- predict(logistic_model, testdata, type = "response")
head(train_predict)
```

```{r}

# change output to probabilities
predicted_probs <- ifelse(train_predict >0.5, 1, 0)
head(predicted_probs)
```
```{r}

table(predicted_probs)
table(testdata$default)
```

